---
title: "Motor Trend Analysis"
author: "Victor Wildner"
date: "7/3/2020"
output: github_document
---

## Question: Do manual vehicles do significantly more miles per gallon than automatic ones?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r message=FALSE}
require(tidyverse)
require(datasets)
require(car)
```

## EDA
```{r cars}
df <- mtcars
```

Transmission type: column `am`:

 - 0 = automatic
 
 - 1 = manual

```{r}
df$am <-  as.factor(df$am)
levels(df$am) <- c("auto", "manual")

fdf <- df %>% 
    select(mpg, am)

df1 <- fdf %>% 
    group_by(am) %>% 
    summarise(
        mean = mean(mpg),
        var = var(mpg)
    )

df1
    
ggplot(fdf, aes(x = mpg, group = factor(am))) +
    facet_grid(rows = vars(factor(am))) +
    geom_boxplot()
```

**Automatic vehicles distribution mean is `r round(df1[2,2] - df1[1,2], 2)` mpg lesser than that of manual vehicles.**

Let's execute t-test to compare the means from both distributions

```{r}
auto <- df[df$am == "auto",]
man <- df[df$am == "manual",]

# Null hypothesis mu_auto = mu_man
pval <- t.test(df$mpg ~ df$am, alternative = 'two.sided', conf.level = 0.95)$p.value
alpha = 0.05
pval
if (pval < alpha) {
    print("Reject the null hypothesis: The means for both distributions are significantly different")
} else {
    print("Do not reject the null hypothesis")
}
```

```{r}
# Features:
# Fit 0 - include all
fit <- lm(mpg ~ ., data = df)
# Fit 1 - include transmission
fit1 <- lm(mpg ~ factor(am), data = df)
print(c('Manual vehicles:',fit1$coef[1]))
print(c('Automatic vehicles:',fit1$coef[2]))

```

Let's check the variance inflation table

```{r}
vif(fit)
```

Add nested models and compare in a variance table

```{r}
# Fit 2 - include weight
fit2 <- lm(mpg ~ factor(am):wt, data = df)
# Fit 3 - include horsepower
fit3 <- lm(mpg ~ factor(am):wt + factor(am):hp, data = df)
# Fit 4 - include cylinders
fit4 <- lm(mpg ~ factor(am):wt + factor(am):hp +  + factor(am):cyl, data = df)

anova(fit1, fit2, fit3, fit4)
```

Adding cylinders feature reduces too little Residual Sum of Squares in the ANOVA variance table and inflates the p-value and F-test outcome, so we remove the cyl feature.

```{r}
# Fit 5 - Remove number of cylinders, include 1/4 mile time
fit5 <- lm(mpg ~ factor(am):wt + factor(am):hp + factor(am):qsec, data = df)
summary(fit5)$coef
```

## Summary

The model explains for 90.5% of the total variance, or 88.3% of the adjusted variance.

- When the weight is increased by 1000 lbs, the mpg rate is `-2.24` for automatic and `-7.93` for manual transmission. Meaning that manual transmission is proven to be better for heavier vehicles. 

- Manual vehicles mpg decrease rate (slope) is less inclined. Meaning that lighter and less weighted vehicles do more miles per gallon.

```{r}
ggplot(df, aes(x = mpg, y = wt, group = am, color = am)) +
    geom_point() +
    geom_smooth(formula = y ~ x, method = 'lm', se = FALSE) +
    theme_minimal()
```